ODGI: scalable tools for pangenome graphs

### 2021-05-31 manuscript/main.tex
Remove table - we'll do this in text
Proposed feature layout

\newcommand{\odgi}{ODGI}
\newcommand{\cmd}[1]{{\texttt{#1}}}
\newcommand{\Rplus}{\protect\hspace{-.1em}\protect\raisebox{.35ex}{\smaller{\smaller\textbf{+}}}}
\newcommand{\pp}{\Rplus\Rplus}
\newcommand{\Cpp}{\mbox{C\Rplus\Rplus}\xspace}

@@

    \subsection{Functionalities}

    Variation graphs are commonly represented by a subset of version 1 of the Graphical Fragment Assembly format
    (\citep{GFAv1}): \texttt{odgi build} and \texttt{odgi view} allow to convert graphs from GFA to ODGI format and
    vice versa, respectively.

    In the variation graph model, paths have to respect the graph's topology: this can be verified with
    \texttt{odgi validate}, to ensure no errors in the input or edited graphs.

    By visualizing pangenome graphs, we can gain insight into the relationship between genomes. \texttt{odgi viz}
    applies binding and direct rendering to a raster image to generate a 1D representation able to scale up to gigabase
    pangenome graphs (Figure~1.A\vphantom{\ref{fig:01}}). However, pangenome graphs may present complex structures to
    visualize: \texttt{odgi sort} finds the best node order to simplify 1D graph visualization by applying several
    cutting-edge sorting algorithms. \texttt{odgi drawn} extends the visualization in 2D, taking in input the layouts
    built by \texttt{odgi layout}.

    Applying \texttt{odgi stats}, users can retrieve metrics describing the graph properties, such as the number of
    nodes, edges, paths, and graph length. ODGI also offers more advanced tools for interrogating the graphs.
    \texttt{odgi bin} summarizes the path information into bins of specified size, enabling a summarized view of
    gigabase scale graphs. \texttt{odgi depth} returns the node depth as defined by query criteria, allowing users to
    detect the complex regions in the graph due to sequences with highly identical repeats. Complex motifs can be also
    detected with \texttt{odgi degree}, which returns the node degree as defined by query criteria. High degree nodes
    are the mirror of misassemblies or problems in the pangenome building, making the tool useful for debugging the
    process as well.

>>>>

\newcommand{\odgi}{ODGI}
\newcommand{\cmd}[1]{{\texttt{#1}}}
\newcommand{\cmdbf}[1]{{\textbf{#1}}}
\newcommand{\topic}[1]{{\cmdbf{#1}}:}
\newcommand{\Rplus}{\protect\hspace{-.1em}\protect\raisebox{.35ex}{\smaller{\smaller\textbf{+}}}}
\newcommand{\pp}{\Rplus\Rplus}
\newcommand{\Cpp}{\mbox{C\Rplus\Rplus}\xspace}

@@

    \subsection{Features}

    \odgi\ has a wide range of commands. Important features are

    \topic{odgi build, view \& validate} pangenome graphs are commonly
    represented by a subset of Graphical Fragment Assembly (GFAv1)
    format\citep{GFAv1}.  \cmd{odgi build} and \cmd{odgi view} allow
    to convert graphs from GFA to ODGI format and vice versa.  In the
    variation graph model, paths have to respect the graph's topology:
    this can be verified with \texttt{odgi validate}, to ensure no
    errors in the input or edited graphs.

    \topic{odgi viz, sort, drawn \& layout} by visualizing pangenome
    graphs, we can gain insight into the relationship between
    genomes. \cmd{odgi viz} applies binding and direct rendering to
    a raster image to generate a 1D representation able to scale up to
    gigabase pangenome graphs
    (Figure~1.A\vphantom{\ref{fig:01}}). However, pangenome graphs may
    present complex structures to visualize: \cmd{odgi sort} finds
    the best node order to simplify 1D graph visualization by applying
    several cutting-edge sorting algorithms. \cmd{odgi drawn}
    extends the visualization in 2D, taking in input the layouts built
    by \cmd{odgi layout}.

    \topic{odgi stats} users can retrieve metrics describing the graph properties, such as the number of
    nodes, edges, paths, and graph length. ODGI also offers more advanced tools for interrogating the graphs.
    \texttt{odgi bin} summarizes the path information into bins of specified size, enabling a summarized view of
    gigabase scale graphs. \texttt{odgi depth} returns the node depth as defined by query criteria, allowing users to
    detect the complex regions in the graph due to sequences with highly identical repeats. Complex motifs can be also
    detected with \texttt{odgi degree}, which returns the node degree as defined by query criteria. High degree nodes
    are the mirror of misassemblies or problems in the pangenome building, making the tool useful for debugging the
    process as well.
